Analysing the Performance of Shallow-Water Models with the Roofline Model

### 2016-01-28 docs/papers/roofline_model/gocean_roofline.tex
roofline: update stream2 results from Archer. Update bandwidth values used in roofline plots to match those in table in text

\begin{table}
\begin{tabular}{|l|c|c|c|c|}
\hline
  \multicolumn{5}{|c|}{E5-1620 v.~2, 3.7GHz} \\
  \hline
LINPACK (GFLOPS) & \multicolumn{4}{c|}{28.32} \\
\hline
 &  L1 Cache & L2 Cache & L3 Cache & Main memory \\
 &   32 KB   & 256 KB   & 10 MB    &  --         \\
\hline
STREAM2 (GB/s)   &  117  &  65.8  & 46.7 &  20.5 \\
  \hline
\multicolumn{5}{|c|}{E5-2697 v.~2, 2.7GHz}  \\
\hline
LINPACK (GFLOPS) & \multicolumn{4}{c|}{25.41} \\
\hline
 & L1 Cache & L2 Cache & L3 Cache & Main memory \\
 &   32 KB  & 256 KB   & 30 MB    &  --         \\
\hline
STREAM2 (GB/s) & 108 & 54.0 & 37.7  &  17.6 \\
\hline
\end{tabular}
\caption{The STREAM2 (DAXPY) and LINPACK benchmark results for a
  single thread running on each of the two CPUs considered here.}
\label{TAB_stream_linpack}
\end{table}

@@

 The equivalent results for the E5-1620
CPU are shown in Figure~\ref{FIG_daxpy_bw}. This plot has the same
shape as that for the E5-2697 but is shifted upwards indicating a
considerable increase in bandwidth accross all levels of the cache
hierarchy. {\bf need to check by re-running on Archer with 200 repeats
  and new min/max results}. In addition, the width of the plateau
corresponding to the L3 cache is narrower because of the smaller size
of this cache on the E5-1620.

@@

Despite the unreliability of FLOP counts, we hope that the trends in
HWPC data as the problem size is varied will be useful. This raw data
is given in Table~\ref{TAB_hwpc_mom_u}.

@@

We measure the mean time to execute the u-Momentum kernel over the
whole domain as 0.0021 seconds (Intel v.16 on the E5-1620) for the
$256^2$ problem size. This gives us a performance of
$\frac{120\times 256^2}{0.0021} = 3.7$ GFLOPS. In contrast, instrumenting
this section of code with the LIKWID marker API and using
likwid-perfctr we measure 4.7 GFLOPS.
For the $128^2$ domain, the measured time is 0.556E-03 s. This gives a
performance of $\frac{120\times 128^2}{0.556\times10^{-3}} = 3.5$
GFLOPS which is some 25\% less than we measure directly with likwid.

@@

 Two performance points are plotted for the Shallow kernel
 corresponding to domains of $512^2$ and $1024^2$. Using the equation
 for working-set size in Table~\ref{TAB_kernel_details}, these
 correspond to 14.7 and 58.7~MB, respectively. Since the E5-2697 has
 an L3 cache of 30~MB, the $512^2$ case should be contained in cache
 while the $1024^2$ case will spill to main memory. This is borne out
 by the plot; the $1024^2$ case is achieving a performance in-line
 with saturating the bandwidth to main memory. The $512^2$ case is
 approaching the L3 bandwidth limit. We attribute the shortfall to the
 fact that the kernel is not run in isolation but as part of a
 time-stepping loop -- other kernels requiring different arrays are
 run on either side of it. {\bf BUT, the working-set for the
   time-stepping loop as a whole has 13 arrays giving a size of 27~MB
   so it should fit in L3 cache?}

In contrast to those results, the performance of the NEMOLite2D
u-Momentum kernel shows little sensitivity to working-set size -- the
$128^2$, $256^2$ and $512^2$ performance points in
Figure~\ref{FIG_roofline_archer} are very close together.

In Figure~\ref{FIG_roofline_haswell} we plot the Roofline model
obtained for the E5-1620 CPU. In order to investigate the poor
performance of the u-Momentum kernel of NEMOLite2D, we also show the
performance of the (simpler) Continuity kernel from that code. As we
would expect, the performance of the u-Momentum kernel remains poor on
this CPU. However, the Continuity kernel performs well. The
working-set sizes for this kernel for the $128^2$ and $256^2$ domains
are approximately 1~MB and 4.2~MB, respectively. Both of these should
therefore fit comfortably in L3 cache. However, again, this kernel is
one of several within the time-stepping loop of NEMOLite2D. For the
$128^2$ domain this loop as a whole has a working-set size of
$18\times128^2\times 8$ bytes or 2.4~MB. {\bf Again, this should all
  fit comfortably in l3 cache so why are we nowhere near the L3
  bandwidth limit? Need to look at performance variation with problem
  size and maybe HWPCs at the same time. Is my figure for l3 bandwidth
  correct? What does loop1 of Shallow achieve on smaller problem
  sizes?}

>>>>

\begin{table}
\begin{tabular}{|l|c|c|c|c|}
\hline
  \multicolumn{5}{|c|}{E5-1620 v.~2, 3.7GHz} \\
  \hline
LINPACK (GFLOPS) & \multicolumn{4}{c|}{28.32} \\
\hline
                 &  L1 Cache & L2 Cache & L3 Cache & Main memory \\
                 &   32 KB   & 256 KB   & 10 MB    &  --         \\
\hline
STREAM2 (GB/s)   &  160  &  61.0  & 46.0 &  20.5 \\
  \hline
\multicolumn{5}{|c|}{E5-2697 v.~2, 2.7GHz}  \\
\hline
LINPACK (GFLOPS) & \multicolumn{4}{c|}{25.41} \\
\hline
               & L1 Cache & L2 Cache & L3 Cache & Main memory \\
               &   32 KB  & 256 KB   & 30 MB    &  --         \\
\hline
STREAM2 (GB/s) & 132      & 59.0     & 40.0     &  20.2 \\
\hline
\end{tabular}
\caption{The STREAM2 (DAXPY) and LINPACK benchmark results for a
  single thread running on each of the two CPUs considered here.}
\label{TAB_stream_linpack}
\end{table}

@@

 The equivalent results for the E5-1620 CPU are shown in
 Figure~\ref{FIG_daxpy_bw}. This plot has the same shape as that for
 the E5-2697 but is shifted upwards indicating an increase in
 bandwidth accross all levels of the cache hierarchy, but particularly
 for L3 and L1. In addition, the width of the plateau corresponding to
 the L3 cache is narrower because of the smaller size of this cache on
 the E5-1620.

@@

Despite the unreliability of the absolute values of the FLOP counts,
we hope that the trends in HWPC data as the problem size is varied
will be useful. This raw data is given in Table~\ref{TAB_hwpc_mom_u}.

@@

We measure the mean time to execute the u-Momentum kernel over the
whole domain as 0.0021 seconds (Intel v.16 on the E5-1620) for the
$256^2$ problem size. This gives us a performance of $\frac{120\times
  256^2}{0.0021} = 3.7$ GFLOPS. (In contrast, instrumenting this
section of code with the LIKWID marker API and using likwid-perfctr we
measure 4.7 GFLOPS.)  For the $128^2$ domain, the measured time is
0.556E-03 s. This gives a performance of $\frac{120\times
  128^2}{0.556\times10^{-3}} = 3.5$ GFLOPS which again is some 25\% less
than we measure directly with likwid.

@@

Two performance points are plotted for the Shallow kernel
corresponding to domains of $512^2$ and $1024^2$. Using the equation
for working-set size in Table~\ref{TAB_kernel_details}, these
correspond to 14.7 and 58.7~MB, respectively. Since the E5-2697 has
an L3 cache of 30~MB, the $512^2$ case should be contained in cache
while the $1024^2$ case will spill to main memory. This is borne out
by the plot; the $1024^2$ case is achieving a performance in-line
with saturating the bandwidth to main memory. The $512^2$ case is
approaching the L3 bandwidth limit. We investigate the reasons for
the fact that it does not achieve the L3 bandwidth limit below.

In contrast to the results obtained with the Shallow kernel, the
performance of the NEMOLite2D u-Momentum kernel shows little
sensitivity to working-set size -- the $128^2$, $256^2$ and $512^2$
performance points in Figure~\ref{FIG_roofline_archer} are very close
together.

In Figure~\ref{FIG_roofline_haswell} we plot the Roofline model
obtained for the E5-1620 CPU. In order to investigate the poor
performance of the u-Momentum kernel of NEMOLite2D, we also show the
performance of the (simpler) Continuity kernel from that code. As we
would expect, the performance of the u-Momentum kernel remains poor on
this CPU. For the $512^2$ domain the Continuity kernel performs well,
reaching the limit imposed by the bandwidth to main memory. However,
the working-set sizes for this kernel for the $128^2$ and $256^2$
domains are approximately 1~MB and 4.2~MB, respectively and therefore
both of these should fit comfortably within L3 cache. Despite this,
the performance achieved with these domain sizes falls short of the L3
bandwidth limit.  Although this kernel is one of several within the
time-stepping loop of NEMOLite2D, the working-set size of this loop
for the $128^2$ domain is just 2.4~MB which again is well within the
capacity of L3.  . Need to look at performance variation with problem
size and maybe HWPCs at the same time.  What does loop1 of Shallow
achieve on smaller problem sizes?
